---
title: "Illustrtion of exponential distribution"
author: "Sayef Ishauqe"
date: "October 08, 2017"
output: html_document
---

Title (give an appropriate title) and Author Name

## Overview: 
In a few (2-3) sentences explain what is going to be reported on.

## Simulations: 
Include English explanations of the simulations you ran, with the accompanying R code. Your explanations should make clear what the R code accomplishes.

* Generate sample data of exponential distribution

```{r }

n<-40
sim<-1000
lamda<-0.2
set.seed(1000)
x<-rexp(n,lamda)

ms<-mean(x);ms
sds<-sd(x);sds
vars<-var(x)

hist(x,breaks = 10, main=paste("Sample mean:",round(ms,2)," Standard Deviation:",round(sds,2)))




```

* Generate simulation data

```{r}
simData<-as.matrix(sample(x,n*sim,replacement=T),sim,n)
means<-apply(simData,1,mean)
vars<-apply(simData,1,var)

par(c(1,2))

hist(means)
hist(vars)

```

Sample Mean versus Theoretical Mean: Include figures with titles. In the figures, highlight the means you are comparing. Include text that explains the figures and what is shown on them, and provides appropriate numbers.
Sample Variance versus Theoretical Variance: Include figures (output from R) with titles. Highlight the variances you are comparing. Include text that explains your understanding of the differences of the variances.
Distribution: Via figures and text, explain how one can tell the distribution is approximately normal.
