---
title: "Analysis of exponential distribution"
author: "Sayef Ishauqe"
date: "October 08, 2017"
output: 
  html_document: 
    fig_height: 3
---

### Overview: 
This assignment aims to analyze a small sample of exponentials to determine its
distribution. Primarily, computer simulation and resampling techniques are 
utilized to determine and compare the characterstics with normal distribution.

### Simulations: 
* Step 1: Generate sample data of exponential distribution
* Step 2: Examine the distribution of the sample data
```{r, echo=TRUE}
n<-40;lambda<-0.2;set.seed(1000)
x<-rexp(n,lambda)
ms<-mean(x); sds<-sd(x); vars<-var(x)
# hist(x,breaks = 10, 
#      main="Distribution of 40 exponetials (sample)")
# text(x=10,y=10,paste("Sample mean:",round(ms,2),"\nStandard Deviation:",round(sds,2)))
```

* Step 3: Generate simulation data (1000 simulations, 40 exponentials)
     + Calculate mean and variance for each simulation
     + Plot histogram for mean and variance

```{r}
sim<-1000
simData<-matrix(sample(x,n*sim,replace=T),sim,n)
means<-apply(simData,1,mean); vars<-apply(simData,1,var); sds<-apply(simData,1,sd)
par(mfrow=c(1,2))
hist(means,breaks = 20,
     main=paste("Dist.of sample mean"),
     xlab="Sample Means")
abline(v=mean(means),lty=5,col="blue")
abline(v=1/lambda,lty=5,col="red")
legend("topright", c(paste("S. Mean:",round(mean(means),2)),paste("T.Mean:",1/lambda)),
       cex=0.55, col=c("blue","red"), lty=5, lwd=2, bty="n");
hist(vars,breaks = 20,xlab="Sample Variances",
     main=paste("Dist. of sample Variance"))
abline(v=mean(vars),lty=5,col="blue")
text(38,100,cex=.8, paste("Sample & pop.\n variance: ~",round(mean(vars),0)))
abline(v=(1/lambda)^2,lty=5,col="red")

```

**Findings - After 1000 simulation, average of sample means are very close to ** 
**Theritical mean. The sample variance is also very close to theoritical** 
**population variance.**

* Step 4: Compare the distribution of 1000 sample means (of 40 random 
exponentials)
     + Draw and compare it with 1000 random exponentials
     + Draw a reference line of nomal distribution
```{r, fig.height=4, fig.width=6}
set.seed(2000)
xx<-rexp(sim,lambda)
hist(means, col=rgb(0,0,1,0.5), breaks=30,xlim=c(-.1,15),ylim=c(0,100),
     main="Distribution of exponentials and its sample means",
     xlab="x",
     ylab="Fequency")
hist(xx,col=rgb(1,0,0,0.5),alpha=1,add=T,breaks = 131)
curve(1000*dnorm(x,mean=1/lambda,sd=1/lambda), lty=2,lwd=3,
      col="gray",from=-1,to=15, n=100,add=TRUE)
legend("topright", c("Dist. of sample means","Dist of 1000 exponentials","Normal Distribution"),
       cex=0.8, col=c("blue","red","gray"), lty=5, lwd=4, bty="n")
```

**Findings - The Distribution of sample means are comparable to normal**
**Distribution curve. However, sample means are more concentrated at the center.**
**So, it shows that exponential distribution holds the characterstrics of normal distribution**