---
title: "Is an automatic or manual transmission better for MPG"
author: "Sayef Ishauqe"
date: "October 21, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Executive Summary

This paper looks into the R standard data `mtcars' to answer the question 
**Is an automatic or manual transmission better for MPG?** and if so, 
**what is difference in MPG for automatic vs manual transmission**

### Method

The analysis concentrates on identifying the best possible multi-variable regression 
model to calcualte mpg and then interpret the infulence of transmission being
manual or automatic in the model.

**Analyze the correlation of variables**

* A ggpairs chart has been generated (***Figure 1***) to see the correlation of 
variables with mpg.
     - `mpg` has high correlation with weight (`wt`), cylinders (`cyl`) and displacement (`disp`)
     - `cyl` and `disp` has high correlation within them. Considering the audience of the magazine `cyl` could be a simiplar predictor to communicate
     - Trasmission, automated or manual (`am`) seems doesn't have high coorelation with `mpg`. 
     - So, developing a multi-variable regression model with minimum number of variable is the best way to move. Once, the model is achieved transmission variable can be introduced to see the impact (for inference / hypothesis).

**Construct and compare models**

* Fit a linear regression model with combination of variables as identified above
     - Test the model with log of wight as well to see it is a better fit. This 
     is because it is not logical to calculate mpg for a car with 0 wight.
     
```{r, echo=TRUE, message=FALSE, warning=FALSE}
fit1<-lm(mpg~wt,data=mtcars)
fit2<-lm(mpg~wt+factor(am),data=mtcars)
fit3<-lm(mpg~wt+factor(cyl),data=mtcars)
fit4<-lm(mpg~wt+factor(am)+factor(cyl),data=mtcars)
fit5<-lm(mpg~I(log(wt))+factor(am)+factor(cyl),data=mtcars)
anova(fit1,fit2,fit3,fit4,fit5)
```

**The model**

**mpg =   37.02 -12.33 x log(wt) 
     - [0.7 for manual transmission] 
     - [3.03 for 6 cyl or, 4.91 8 cyl]**

```{r, echo=FALSE}
summary(fit5)$coef; confint(fit5)
```

* The model also indicates that automatic transmission vehicles (`am==0`) 
produces slightly higher mpg (0.7) than that of manual transmission vehicle.
However, Further looking into the confidence interval of the cofficients, the 
95% interval of transimssion variable `am` included 0. So, the hypothesis that 
this variable has no impact on milage can not be rejected.


## Conclusion
The model shows that millage is mostly correlated with weight and than number of 
cylinders of a car. 

```{r, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE}
library(ggplot2)
g<-ggplot(data=mtcars,aes(y=mpg,x=log(wt),fill=factor(am)))
g<-g+geom_point(col="red")
# g<-g+facet_grid(cyl~.)
g<-g+scale_fill_discrete(name="Transmission Type",labels=c("Automatic","Manual"))
g<-g+geom_text(aes(label=cyl),nudge_y = 1)
g<-g+ggtitle(label="Model showing correlation of mpg with log(wt), cyl & transmission type")
g<-g+geom_smooth(method="lm")
g
```

While in preliminary anaysis, it showed that manual transmission
provides higher mpg (***Figure 2***), the model has produced high residuals. 
A better fit model shows that transmission type doesn't have significant impact on 
mpg. Analysis of residuals is shown in ***Figure 3***. The suggested model is still not safe
from high residuals which can be further analyzed to refine the multi-variable 
regressional model.


## Appendix

```{r, echo=FALSE, fig.height=6, fig.width=7, message=FALSE, warning=FALSE}
#install.packages("GGally")
require(GGally)
ggpairs(mtcars,lower=list(contenuous="smooth"))
```
***Figure 1: Pairs plot of all variables in mtcars***

```{r, echo=FALSE}
fit<-lm(mpg~factor(am),data=mtcars)

summary(fit)

```
***Figure 2: correlation of transmission type with mpg***

```{r, echo=FALSE}
par(mfrow=c(2,2))
plot(fit5)
```
***Figure 3: Residuals of the suggested model***

