---
title: "The response of supplements on Tooth Growth of Guinea Pigs"
author: "Sayef Ishauqe"
date: "October 09, 2017"
---

### Overview
This papaer analyzes the Tooth Growth data available in R datasets to infer the 
response of supplements on tooth growth from the sample. The analyis will 
evaluate hypothesises across delivery methods (Orange Juice - OJ, 
ascrobic acid - VC) and dose levels (0.5,1,2 mg/day). 

### Data Processing & Analysis
* Step 1: Load the data and examine the structure
```{r, echo=FALSE}

tData<-ToothGrowth
# head(tData)
levels(tData$supp)<-c("Orange Juice","Ascrobic Acid (Vitamin C)")
tData$dose<-factor(tData$dose)
# colnames(tData)<-c("Tooth.length", "supplement.Type","Dose.in.mg.per.day")
summary(tData)
```
* Step 2: Plot a bar chart of average tooth growth for different supplements and 
dose levels

```{r, include=FALSE}
library(dplyr)
```

```{r, echo=FALSE, fig.height=3.5}

tDataGrouped<-group_by(tData,supp,dose)
tDataSummary<-summarise_all(tDataGrouped,mean)
library(ggplot2)
g<-ggplot(data=tDataSummary,aes(y=len,x=factor(dose),fill=supp))
g<-g+geom_col(position = "dodge")
g<-g+geom_text(aes(label=len), nudge_y = 1,nudge_x=c(-.2,-.2,-.2,.2,.2,.2))
# g<-g+coord_flip()
g<-g+ scale_fill_discrete(name="supplement Type")
g<-g+theme(legend.position=c(.15,.85))
g<-g+labs(x="Dose (mg/day)", y="Average Teeth length")
g<-g+ggtitle(label="Higher supplement dose can be related to higher tooth growth") 
print(g)
```

**Findings - Higher average teeth lenght is observed for higher doses.**

* Step 3: Perform t-tests to validate the hypothesis that higher doses is related 
with tooth growth
     + Conduct T-test between dose level 2 and 0.5
     
          + H0-Dose levels is not related with tooth growth
          + H1- 2 Mg/day is related with higher tooth growth than 1 mg/day dose
          + H2- 1 Mg/day is related with higher tooth growth than 1 mg/day dose
     
          *The experiment was run on 60 pigs with different dose levels and supplements.*
          *Same pig was not administered with different dose levels for this expericemnt*
          *So, its not a paired sample. However, the population variance of pig tooth*
          *lenght can be assumed to be identical*


```{r}
tDataDose12<-tData[tData$dose!=1.0,]
tres<-t.test(len~dose,data=tDataDose12,paired=F,var.equal=T)

```

**Findings - The null hypothesis (H0) can be rejected as p-value(`r tres$p.value`) is very small**

* Step 4: Conduct a power t-test between 2 and 1 mg/day and 1 and .5 mg/day

```{r, include=FALSE}
meanD.5<-mean(tData[tData$dose==.5,]$len)
meanD1<-mean(tData[tData$dose==1,]$len)
meanD2<-mean(tData[tData$dose==2.0,]$len)

sdD.5<-sd(tData[tData$dose==.5,]$len)
sdD1<-sd(tData[tData$dose==1,]$len)
sdD2<-sd(tData[tData$dose==2.0,]$len)

```

```{r}
pt1<-power.t.test(n=20, delta = meanD2-meanD1, sd = sqrt((sdD2^2+sdD1^2)/2), 
             sig.level = 0.05,
             type = "two.sample",
             alternative = "one.sided")

pt2<-power.t.test(n=20, delta = meanD1-meanD.5, sd = sqrt((sdD.5^2+sdD1^2)/2), 
             sig.level = 0.05,
             type = "two.sample",
             alternative = "one.sided")
pt2
```

**Findings - The sample data provide enough (high power: `r round(pt1$power,4)`,`r round(pt2$power,4)`)**
**evidence to support the hypothesises**
**that higher doses of supplements can be related to higher tooth growth**

* Step 5: Perform t-tests to validate the hypothesis that supplements are related to
different tooth growth
     + H0-With similar dose levels, different supplements are not related to different
     tooth growth
     + H1- Orange Juice (OJ) is a better supplement related with tooth growth
          + Conduct T-test between supplements 

```{r}
tres<-t.test(len~supp,data=tData,paired=F,var.equal=T)
```
***Findings - The hypothesis (H0) that different supplements does not produce***
***different result (difference in mean) - can not be rejected(p-value `r round(tres$p.value,2)`)***

* Step 6:Conduct a power test to evaluate the number of samples required to evaluate
which supplement is better
     
```{r, echo=FALSE}
meanOJ<-mean(tData[tData$supp=="Orange Juice",]$len)
meanVC<-mean(tData[tData$supp=="Ascrobic Acid (Vitamin C)",]$len)

sdOJ<-sd(tData[tData$supp=="Orange Juice",]$len)
sdVC<-sd(tData[tData$supp=="Ascrobic Acid (Vitamin C)",]$len)

power.t.test(power = .90, delta = meanOJ-meanVC, sd = sqrt((sdOJ^2+sdVC^2)/2), 
             sig.level = 0.05,
             type = "two.sample",
             alternative = "one.sided")
```
***At least 71 (meaning total 142) pigs has to administered with alternative supplements***
***to conclude whether Orange Juice (Supp-OJ) is better supplement than Ascrobic Acid***
***(Supp-VC) for tooth growth***

### Conclusion
supplements are related to higher tooth growth of pigs. However, the sample size
is not large enough to conclude which supplement is more effective.